---
title: 'Lab 1: Introduction to R'
author: "Caitlyn Ralph"
date: "2/7/2018"
output: 
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic Commands

* Functions perform operations: `funcname(input1, input2)`
* Create a vector of numbers:
```{r}
x <- c(1,3,2,5)
x
```
* Open new help file window: `?funcname`
* Add two sets of numbers together:
```{r}
x <- c(1,6,2)
y <- c(1,4,3)
x
y
length(x)
length(y)
x + y
```
* Look at a list of all objects: `ls()`
* Delete any that we don't want: `rm()`
* Remove all objects at once: `rm(list=ls())`
* Matrices:
```{r}
x=matrix(data=c(1,2,3,4), nrow=2,ncol=2)
x
x=matrix(c(1,2,3,4),2,2)
x
matrix(c(1,2,3,4),2,2,byrow=TRUE)
sqrt(x)
x^2
```
* `r(norm)` generates a vector of random normal variables, with first argument n the sample size:
``` {r}
x=rnorm(50)
y=x+rnorm(50,mean=50,sd=.1)
cor(x,y)
```
* `set.seed()` allows our code to reproduce exact same set of random numbers
  + Used whenever calculations involve random quanities
* `mean()` and `var()` (`sqrt(var())`=standard deviation):
```{r}
set.seed(3)
y=rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)
```

## Graphics

* Primary way to plot data: `plot()`
  + Scatterplot of the numbers in x versus the numbers in y: `plot(x,y)`
  + Additional options, e.g. `xlab` applies label on the x-axis
```{r}
x=rnorm(100)
y=rnorm(100)
plot(x,y)
plot(x,y,xlab="this is the x-axis",ylab="this is the y-axis",main="Plot of X vs Y")
```

* Save output of an R plot: `pdf()`, `jpeg()`, etc
* Create a sequence of numbers: `seq(a,b)`
```{r}
x=seq(1,10)
x
x=1:10
x
x=seq(-pi,pi,length=50)
```
* Represent three-dimensional data: Contour plot
  + Takes three arguments:
    1. A vector of the x values (first dimension)
    2. A vector of the y values (second dimension)
    3. A matric whose elements correspond to the z value (third dimension) for each pair of (x,y) coordinates
```{r}
y=x
f=outer(x,y,function(x,y)cos(y)/(1+x^2))
contour(x,y,f)
contour(x,y,f,nlevels=45,add=T)
fa=(f-t(f))/2
contour(x,y,fa,nlevels=15)
```

* `image()`: Works the same as `contour()`, except it produces a color-coded plot whose colors depend on the z value (AKA a heatmap)
* `persp()`: Produce a three-dimensional plot
  + `theta` and `phi` control the angels at which the plot is viewed
```{r}
image(x,y,fa)
persp(x,y,fa)
persp(x,y,fa,theta=30)
persp(x,y,fa,theta=30,phi=20)
persp(x,y,fa,theta=30,phi=70)
persp(x,y,fa,theta=30,phi=40)
```

## Indexing Data

```{r}
A=matrix(1:16,4,4)
A
A[2,3]
A[c(1,3),c(2,4)]
A[1:3,2:4]
A[1:2,]
A[,1:2]
A[1,]
A[-c(1,3),]
A[-c(1,3),-c(1,3,4)]
dim(A) # number of rows followed by the number of columns given a matrix
```

## Loading Data

* `read.table()` imports a data set
* `write.table()` exports data
```{r}
Auto=read.table("Auto.data.txt",header=T,na.strings="?")
#fix(Auto)
```
* `read.csv`
```{r}
Auto=na.omit(Auto)
dim(Auto)
```
```{r(}
names(Auto)
```

## Additional Graphical and Numerical Summaries

```{r}
plot(Auto$cylinders, Auto$mpg)
attach(Auto)
plot(cylinders, mpg)
cylinders=as.factor(cylinders)
plot(cylinders, mpg)
plot(cylinders, mpg, col="red")
plot(cylinders, mpg, col="red", varwidth=T)
plot(cylinders, mpg, col="red", varwidth=T,horizontal=T)
plot(cylinders, mpg, col="red", varwidth=T, xlab="cylinders", ylab="MPG")
```
```{r}
hist(mpg)
hist(mpg,col=2)
hist(mpg,col=2,breaks=15)
```
```{r}
pairs(Auto)
pairs(~ mpg + displacement + horsepower + weight+ acceleration, Auto)
```
```{r}
plot(horsepower,mpg)
identify(horsepower,mpg,name)
```
```{r}
summary(Auto)
summary(mpg)
```
